The importance of a complete statistical description of dependence between variables for ecology and related fields

### 2019-05-02 Paper.Rmd

<!--Discuss stuff here on tradeoff and life history that Bever told me about-->
The relationships we studied between BMR and body mass relate
to a trade-off between mass-specific BMR (i.e., BMR per unit body mass) and body
mass itself. Trade-offs between life history characteristics are a focal topic in 
evolutionary ecology. There are reasons to believe copulas interrelate with 
life-history theory in additional ways beyond what we could demonstrate 
with the data we used.
It is well known that energy allocation to a life function, X (e.g., reproduction)
will reduce the energy that can be allocated to other functions, $Y_1$, $Y_2$, $Y_3$ 
(e.g., growth, predation avoidance; this is sometimes called the principle of 
allocation).
But it is equally well known that X can trade-off against any or all of 
the $Y_i$. Therefore, for large values of X which approach absolute 
limitations, there may be a strong 
association between X and $Y_1$. Whereas for small values of X there may be
little association - resources not allocated to X can instead be allocated 
to $Y_1$ or $Y_2$ or to some combination. This constitutes 
asymmetric tail dependence between X and $Y_1$. For instance, \cite{Winemiller1992}
described a three-way trade-off in North American fishes between age of reproductive
maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
produce a tight association between age of maturity and fecundity for 
fishes with low age of maturity, but little association between those
variables for later-maturing
fishes because those fishes may invest much or little in juvenile survival.
Although copulas may be very useful for studying multi-dimensional trade-offs, 
applications will 
require careful attention to the possible consequences of biased sampling:
if the degree of completeness of a dataset is associated with one or
more of the quantities being studied then statistical artefacts
may bias conclusions of copula analyses (Appendix \ref{SM-sect:missingdata}, 
Fig. \ref{SM-fig:missingdata}).

>>>>

<!--Discuss stuff here on tradeoff and life history that Bever told me about-->
The relationships we studied between BMR and body mass relate
to a trade-off between mass-specific BMR (i.e., BMR per unit body mass) and body
mass itself. Trade-offs between life history characteristics are a focal topic in 
evolutionary ecology. There are reasons to believe copulas interrelate with 
life-history theory in additional ways beyond what we could demonstrate 
with the data we used.
It is well known that energy allocation to a life function, X (e.g., reproduction)
will reduce the energy that can be allocated to other functions, $Y_1$, $Y_2$, $Y_3$ 
(e.g., growth, predation avoidance; this is sometimes called the principle of 
allocation).
But it is equally well known that X can trade-off against any or all of 
the $Y_i$. Therefore, for large values of X which approach absolute 
limitations, there may be a strong 
association between X and $Y_1$. Whereas for small values of X there may be
little association - resources not allocated to X can instead be allocated 
to $Y_1$ or $Y_2$ or to some combination. This constitutes 
asymmetric tail dependence between X and $Y_1$. For instance, @Winemiller1992
described a three-way trade-off in North American fishes between age of reproductive
maturity, juvenile survivorship, and fecundity. The trade-off should, in theory,
produce a tight association between age of maturity and fecundity for 
fishes with low age of maturity, but little association between those
variables for later-maturing
fishes because those fishes may invest much or little in juvenile survival.
Although copulas may be very useful for studying multi-dimensional trade-offs, 
applications will 
require careful attention to the possible consequences of biased sampling:
if the degree of completeness of a dataset is associated with one or
more of the quantities being studied then statistical artefacts
may bias conclusions of copula analyses (Appendix \ref{SM-sect:missingdata}, 
Fig. \ref{SM-fig:missingdata}).
